---
title: "ESM244 Code-Along 1"
author: "Taylor Lockmann"
date: "1/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(sf)
library(tmap)
```

### Read in SF trees data

```{r}
sf_trees <- read_csv(here("data", "sf_trees", "sf_trees.csv"))

```

### Basic wrangling reminders

Refresh some skills for data wrangling & summary statistics using functions in the `dplyr` package.

Find the top 5 highest observations of trees by legal_status, do some wrangling, make a graph.
```{r}
top_5_status <- sf_trees %>% 
  count(legal_status) %>% 
  drop_na(legal_status) %>% 
  rename(tree_count = n) %>% 
  relocate(tree_count) %>% 
  slice_max(tree_count, n = 5)
```

Make a graph of those top 5 observations by legal status.

```{r}
ggplot(data = top_5_status, aes(x = fct_reorder(legal_status, tree_count), y = tree_count))+
  geom_col()+
  labs(x = "Legal Status", y = "Tree Count")+
  coord_flip()+
  theme_minimal()
```

### A few more data wrangling refresher examples

Only want to keep observations (rows) for Blackwood Acacia trees

```{r}
blackwood_acacia <- sf_trees %>%
  filter(str_detect(species, "Blackwood Acacia")) %>% 
  select(legal_status, date, latitude, longitude)

ggplot(data = blackwood_acacia, aes(x = longitude, y = latitude))+
  geom_point()
```

### tidyr::separate() and unite() functions

Useful for combining or separating columns

```{r}
sf_trees_sep <- sf_trees %>% 
  separate(species, into = c("spp_scientific", "spp_common"), sep = "::")
```

Example: tidyr::unite() 

```{r}
sf_trees_unite <- sf_trees %>% 
  unite("id_status", tree_id:legal_status, sep = "_cool!_")
```

